Decision Assistance and Steering in Health Insurance

### 2021-06-14 finals/paper.tex

\begin{document}
\title{Decision Assistance and Steering in Health Insurance}
\author{Ian McCarthy \& Evan Saltzman}
\date{May 2021}
\maketitle

\vspace{-2ex}
\begin{abstract}
\noindent 
\end{abstract}

@@

Second, we consider the potential for decision assistance to ``improve'' plan choice. While identifying ``improvement'' is not generally feasible in our data, we can identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set. For example, Gold and Platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150\% of the federal poverty level. We can therefore identify dominated choices and examine whether those with decision assistance appear less likely to select a dominated plan. As with our initial nested logit model of plan choice, our estimand is the ATT for households receiving decision assistance, and our estimator is an augmented inverse propensity weighted logit regression.

Regarding dominated choices, our results suggest that individuals with decision assistance are over 1 percentage point less likely to make a dominated choice. On a base of 3.5\%, this reflects a 30\% decrease in the probability of making a dominated choice. We again find evidence of heterogeneities in the effects of different forms of decision assistance. In particular, we find that individuals using publicly provided decision assistance are around 0.76 percentage points less likely to select a dominated plan, while individuals using a private insurance agent are 1.2 percentage points less likely to select a dominated plan.

@@

\subsection{Decision Assistance}
There are two forms of formal decision assistance available to potential enrollees on the California exchange: publicly funded ``navigators;'' and private insurance agents or brokers. For states participating in the federal exchange, navigators are mandated under the ACA and federally funded. Navigators are voluntary for state-based exchanges, although most states offer such assistance to potential enrollees, including California.

A navigator is essentially a public employee available to help potential enrollees better understand their insurance options. Access to navigators has been shown to have a particularly large effect on Medicaid enrollment \citep{myerson2019, sommers2015, aizer2003}. Relatedly, navigators tend to work with households of lower incomes, currently uninsured, and non-native english speakers.

@@

Our final data consist of 8.3 million total annual enrollments from 2014-2019. Of these, 4.5 million are newly enrolled (e.g., not enrolled in any California exchange plan in the prior year). Table~\ref{tab:summary-stats} displays summary statistics for Covered California households, separately between those that used and did not use decision assistance. The silver tier is the most commonly selected option because consumers eligible for cost sharing reductions (CSRs) must choose a silver plan to receive CSRs.  Approximately 68 percent of California enrollees and 61 percent of Washington enrollees are eligible for CSRs, while 91 percent of California enrollees and 85 percent of Washington enrollees are eligible for premium subsidies.  The proportion of consumers exempt from the individual mandate is small, but is notably higher among those who are uninsured.  The uninsured rate is substantially higher among young adults and males in both states. Individuals with incomes between 250 and 400 percent of FPL, who receive relatively small subsidies, and those with incomes above 400 percent of FPL, who are ineligible for subsidies, make up a relatively large share of the uninsured population.

\begin{table}[h]
\centering	
\caption{Summary Statistics for Covered California Enrollees}
\input{tables/summary-stats.tex}
\footnote{Something}
\label{tab:summary-stats}
\end{table}

@@

Interpreting our results as an estimate of the ATT assumes that individuals receiving decision assistance do not seek assistance due to some unobserved factors that also affect their plan choice. Conditional on household characteristics as well as insurer fixed effects, interacted with a full set of region and year fixed effects, we argue that this assumption is plausible.

\section{Dominated Choices}
\label{sec:dominated}
As a complementary empirical analysis, we identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set, and we estimate the effects of decision assistance on the probability of making such a choice. For example, Gold and Platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150\% of the federal poverty level. We estimate the effect of decision assistance on dominated choices with a linear probability model, allowing for year, insurer, and region fixed effects.

@@

\subsection{Model of Commissions and Steering}
\label{subsec:steering-model}

\subsection{Structural Estimation}
\label{subsec:steering-methods}

>>>>

\begin{document}
\title{Decision Assistance and Steering in Health Insurance}
\author{Ian McCarthy \& Evan Saltzman}
\date{June 2021}
\maketitle

\vspace{-2ex}
\begin{abstract}
\noindent In many markets, consumers have access to an agent or some other intermediary to aid in decision making. The influence of such agents on consumer decisions, and the firm's influence on the behaviors of its agents, have important implications for the efficiency of these markets. In this paper, we study the welfare effects of decision assistance and firm steering in health insurance using the population of enrollments from the California Affordable Care Act (ACA) exchange from 2014 to 2019. We find that enrollees are nearly 2x more likely to select a silver plan when using some form of decision assistance and 30\% less likely to select a ``dominated'' plan. Preliminary results suggest that firms have considerable ability to steer consumers' decisions using agent commissions, where a \$1 increase in commissions paid to agents has the same effect as a \$2 or \$3 increase in plan premiums.
\end{abstract}

@@

Second, we consider the potential for decision assistance to ``improve'' plan choice. While identifying ``improvement'' is not generally feasible in our data, we can identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set. For example, gold and platinum tier plans are dominated for any household that is eligible for cost-sharing subsidies and with incomes below 150\% of the federal poverty level. We can therefore identify dominated choices and examine whether those with decision assistance appear less likely to select a dominated plan. As with our initial nested logit model of plan choice, our estimand is the ATT for households receiving decision assistance, and our estimator is an augmented inverse propensity weighted logit regression.

Regarding dominated choices, our results suggest that individuals with decision assistance are over 1 percentage point less likely to make a dominated choice. On a base of 3.5\%, this reflects a 30\% decrease in the probability of making a dominated choice. We again find evidence of heterogeneities in the effects of different forms of decision assistance. In particular, we find that individuals using publicly provided decision assistance are around 0.8 percentage points less likely to select a dominated plan, while individuals using a private insurance agent are 1.2 percentage points less likely to select a dominated plan.

@@

\subsection{Decision Assistance}
There are two forms of formal decision assistance available to potential enrollees on the California exchange: publicly funded ``navigators'' and private insurance agents or brokers. For states participating in the federal exchange, navigators are mandated under the ACA and federally funded. Navigators are voluntary for state-based exchanges, although most states offer such assistance to potential enrollees, including California.

A navigator is essentially a public employee available to help potential enrollees better understand their insurance options. Access to navigators has been shown to have a particularly large effect on Medicaid enrollment \citep{myerson2019, sommers2015, aizer2003}. Relatedly, navigators tend to work with households of lower incomes, currently uninsured, and non-native english speakers. Among individuals receiving some form of assistance in our data, about 30\% of households used a navigator.

Potential enrollees on the California exchange may also use a traditional private insurance agent or broker (hereafter, agents). Agents are not generally tied to any specific insurer and can enroll individuals in any available plans. Agents are also not restricted to any specific geography or region and can enroll individuals in the relevant rating area regardless of the physical office in which the agent resides. Among individusl receiving some form of assistance in our data, 70\% used a private insurance agent.

Unlike navigators, agents are paid commissions for each enrollment into an exchange plan. The amount of commission varies by insurer and is generally set as a flat fee per enrollee or as a percentage of the premium paid by that enrollee. For new enrollees, flat fee commissions typically range between \$100 and \$200, with percent commissions ranging between 2\% and 5\%. Figure~\ref{fig:commissions} presents the observed commission rates by insurer over time, separately among insurers offering a flat fee commission versus those offering a percentage commission. As is evident from the figure, there exists relatively large variation in commissions across insurers, as well as within-variation over time for large insurers such as Anthem, Health Net, and Blue Shield.

@@

Our final data consist of 8.3 million total households from 2014-2019. Of these, 4.5 million were not previously enrolled in any Covered California plan in the prior year. Table~\ref{tab:summary-stats} displays summary statistics for all households in our data. Since the key independent variable throughout much of our analysis is whether a household used some form of decision assistance, we present summary statistics separately between those that used and did not use decision assistance.

We see from Table~\ref{tab:summary-stats} that enrollees using decision assistance are of lower income and larger household size. Among types of plans purchased, we see little difference between those with and without decision assistance, with the exception of Kaiser plans for which individuals using decision assistance are less likely to purchase. Overall, the silver tier is the most commonly selected option, and this is particularly true for those using decision assistance. Enrollees with decision assistance are also much more likely to select an HMO and relatively more likely to select a PPO.

@@

Figure~\ref{fig:ps-balance} presents the resulting propensity scores from our preliminary logit regression of decision assistance against observable household demographics. As is evident from the figure, there exists common support in propensity scores for the treated and control groups. In Figure~\ref{fig:cov-balance}, we present a comparison of observable household characteristics between the treatment and control groups. The hollow circles in Figure~\ref{fig:cov-balance} reflect the unadjusted differences in means, while the solid circles reflect the difference in means when weighting households. Weights for treated households are set to 1, and weights for untreated households are set to  $\frac{\hat{p}_{i}}{1-\hat{p}_{i}}$ based on the preliminary logistic regression. We see from Figure~\ref{fig:cov-balance} that our treated and control groups are well-balanced after weighting.

We summarize our ATT estimates by insurer and metal tier, presented in Figures~\ref{fig:choice-insurer} and \ref{fig:choice-metal} respectively. Figure~\ref{fig:choice-insurer} shows that individuals using decision assistance are significantly more likely to select a plan offered by Blue Shield, Health Net, or another smaller insurer, and much less likely to select a plan offered by Anthem or Kaiser. Of note, Anthem and Kaiser are similar in that they both offer flat commission payments. Kaiser sets the lowest of all such flat commissions in all years, while Anthem dropped their commission payments dramatically beginning in 2015. We investigate the issue of commission and plan choice more fully in Section \ref{sec:steering}.

\section{Dominated Choices}
\label{sec:dominated}
The results in Section~\ref{subsec:causal-results} reveal large effects of decision assistance on plan choice; however, it is not clear from those results whether decision assistance \textit{improves} plan choice or simply reallocates households into different, perhaps worse, plans.

To examine this question, we first identify specific instances in which the observed plan choice is dominated by some other plan in an individual's choice set, and we estimate the effects of decision assistance on the probability of making such a choice. For example, for any household that is eligible for cost-sharing subsidies and with incomes below 150\% of the federal poverty level (FPL), silver plans will have lower premiums and lower out-of-pocket costs compared to gold or platinum tier plans such that the actuarial value of silver plans exceeds that of gold and platinum plans. Since provider networks are not allowed to vary across metal tiers for the same plan type, gold and platinum plans are strictly dominated by silver plans for such households. Gold plans are similarly dominated by silver plans for households with incomes between 150\% and 200\% of the FPL.

We estimate the effect of decision assistance on dominated choices with a logistic regression, in which the outcome is an indicator for whether the enrollee made a dominated choice. Individuals that do not have a dominated plan in their choice set are excluded. Following the process laid out in Section~\ref{sec:causal}, we estimate a weighted logistic regression model only among individuals without any decision assistance, with weights derived from a preliminary logit regression of decision assistance against observable household characteristics. We then use the coefficient estimates to form predicted counterfactual probabilities for dominated choices among individuals that did use decision assistance. Finally, again limiting to those with decision assistance, our estimate of the ATT is the average probability of a dominated choice less the average predicted probability of a dominated choice.

Results are presented in Figure~\ref{fig:dom-choice}, where we show the point estimate and 95\% confidence interval for the ATT. We also report the mean proportion of dominated choices observed in the data. We present different ATTs for people with any decision assistance and for those specifically using public navigators versus private insurance agents. We estimate that individuals are just over 1 percentage point less likely to make a dominated choice when using any form of decision assistance. On a base of 3.5\%, this reflects a nearly 30\% reduction in the prevalence of dominated choices. These effects are largest among those using a private insurance agent, where we estimate an ATT of -1.2 percentage points compared to an ATT of -0.8 percentage points for individuals using a navigator.

@@

\subsection{Model}
\label{subsec:steering-model}

Consider a two-stage  game where  (1) firms set premiums and commissions simultaneously to maximize their expected profit and (2) consumers select a plan to maximize their expected utility. In the following three subsections, we detail how we (1) model the decisions of firms; (2) model the decisions of consumers; and (3) find the model equilibrium.

Given our available data, we do not explicitly model the actions of individual insurance agents.  However, we do model how firms use agent compensation to steer an agent's clients to their plans.  Our approach assumes that the consumer's decision to seek agent assistance is independent of the firms' commission levels.  This identifying assumption is valid if commissions incentivize agents to recommend certain plans to their clients, but do not induce agents to change advertising levels or marketing efforts to  seek new clients. Advertising and marketing by both insurers and agents are tightly regulated, particularly in Covered California.  We therefore believe our identifying assumption is relatively  innocuous.

\subsubsection{Stage 1: Firms Set Premiums and Commissions}

\paragraph{Firm Decision Variables:}
Firms set premiums and commissions to maximize their expected profit.  Let $p_{jmt}$ be the base premium decision for each plan $j$ sold in market $m$ in period $t$ and the \textit{commission decision} $n_{gjt}$

\vspace{-0.4in}		
\singlespacing
\begin{eqnarray*}
    n_{gjt}  \equiv 
    \begin{cases} 
	    \delta_{gjt} & \mathrm{\% \,\, commission} \\		
	    \eta_{gjt} & \mathrm{\$ \,\, commission} \\
	    0 & \mathrm{o/w}
    \end{cases}
\end{eqnarray*}		   
\doublespacing
\noindent for each plan $j$ sold at time $t$ with renewal status $g$ (i.e., new or existing customer). Agents are either paid a percentage $\delta_{gjt}$  of the consumer's unsubsidized premium  or a flat amount $\eta_{gjt}$ that is independent of the premium.  We assume the firm's decision to pay a percentage commission or flat commission is exogenous because switching agent compensation methods is costly.  The only firm that switched between compensation methods during our six-year study timeframe was Health Net following its acquisition by Centene. The \textit{commission paid} to household $i$'s insurance agent is

\vspace{-0.4in}		
\singlespacing			
\begin{eqnarray*}
	\eta_{gijt}(\textit{p},\textit{n})  \equiv 
	\begin{cases} 
		n_{gjt} \left(\sigma_{it}p_{jmt}\right) & \mathrm{\% \,\, commission} \\		
		n_{gjt} & \mathrm{\$ \,\, commission}
    \end{cases},
\end{eqnarray*}
\doublespacing
\noindent where $\textit{p} = {\bf p}_t$ is the vector of plan base premiums set by all insurers in each market in year $t$, $\textit{n} = {\bf n}_t$ is the vector of commission decision variables in period $t$, and the household's (unsubsidized) premium is the product of the household's rating factor $\sigma_{it}$  and the plan's base premium $p_{jmt}$. In California, the rating factor can vary by age and geographic residence of the household's members.\footnote{The ACA also permits rating by tobacco usage, but California prohibits tobacco rating.}  Insurers can charge a 64-year-old up to 3 times as much as a 21-year-old. Figure~\ref{rating_area_partitions} shows the partition of California into rating areas.  In each rating area,  an insurer's premium must be the same for all consumers of the same age. Total commissions paid by firm $f$  are

\vspace{-0.2in}		
\begin{equation*}
    \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) = \sum_{g \in G, i \in I, m\in M, k \in J_{fmt}} \left( \mathbb{I}_{n,i,t} \right)  q_{ikt}(\textit{p},\textit{n};\boldsymbol{\beta}) \eta_{gikt}(\textit{p},\textit{n}),
\end{equation*}
where $\mathbb{I}_{n,i,t}$ indicates whether household $i$ is assisted by an agent in year $t$, $q_{ikt}(\cdot)$ is the probability household $i$ chooses plan $k$ at time $t$, and $J_{fmt}$ is the set of plans offered by firm $f$ in market $m$ at time $t$.  The model parameters $\boldsymbol{\theta} \equiv (\boldsymbol{\beta},\boldsymbol{\gamma},\boldsymbol{\mu})$, where $\boldsymbol{\beta}$ are the demand parameters, $\boldsymbol{\gamma}$ are the risk score parameters, and $\boldsymbol{\mu}$ are the average claims parameters.

\paragraph{Risk Adjustment Transfers:} 
\citet{Pope2014} derive the ACA risk adjustment transfer formula.\footnote{We start with \citet{Pope2014}'s transfer formula as derived in their first appendix, which allows plans to vary only by their actuarial values (and not by differences in firm efficiency, geographic costs, allowable rating factors, or moral hazard).  We start with this formula because we want to capture all differences in expected risk, except for cost sharing and any associated moral hazard, in the plan's risk score (i.e., cost sharing and moral hazard are addressed through the utilization share).  In contrast, the plan liability risk score   $PLRS_j$ as defined in \citet{Pope2014}'s second appendix does not account for certain differences such as variation in geographic cost.  Instead, \citet{Pope2014} account for these differences by applying factors in the transfer formula.}  In our notation, the per-member per-month risk adjustment transfer $ra_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$ for plan $j$ is

\vspace{-0.2in}
\begin{equation*}
	ra_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) = \left(\frac{r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})\sum_{m \in M, l \in J_{mt}} q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}} r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}    -     \frac{h_j\sum_{m \in M, l \in J_{mt}} q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}} h_l q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}\right)\overline{p},
\end{equation*}
where $r_{jmt}(\cdot)$ is the plan risk score, $h_j$ is an exogenous expected utilization factor set by regulation that accounts for the plan AV and any associated moral hazard, and $\overline{p}$ is the weighted average premium in the market.  The plan's total risk adjustment transfer $RA_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$ equals

\vspace{-0.4in}
\begin{eqnarray}
\label{eqn:ra_formula}
	RA_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) &=& ra_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})  q_{jmt}(\textit{p},\textit{n};\boldsymbol{\beta}) \nonumber \\
	&= & \left[rs_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})  - us_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})  \right] R_t(\textit{p},\textit{n};\boldsymbol{\beta}),
\end{eqnarray}
where $rs_{jmt}(\cdot)$ is the plan's ``risk share'' of total claims, $us_{jmt}(\cdot)$ is the plan's ``utilization share'' of total claims, and $R_t(\cdot) = \sum_f R_{ft}(\cdot)$ is total premium revenue across all plans. The total risk adjustment transfer for the firm equals the sum of the risk adjustment transfers for the plans that it sells (i.e., $RA_{ft}(\cdot) = \sum_{m \in M, j \in J_{fmt}} RA_{jmt}(\cdot)$.

The plan's risk share includes the combined effects of adverse selection, moral hazard, and plan AV, whereas the plan's utilization share only includes the effects of moral hazard and plan AV. Thus, the difference between the risk share and utilization share captures the plan's relative risk due to adverse selection only. The risk share equals

\vspace{-0.1in}
\begin{equation*}
	rs_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) = \frac{ r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) q_{jmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}}  r_{lmt}(\textit{p},\textit{n};\boldsymbol{\theta}) q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})},
\end{equation*}
where $J_{mt}$ is the set of all plans offered in market $m$ and year $t$, and the plan risk score $r_{jmt}(\cdot)$ is a function of enrollee characteristics and the plan AV. We do not directly observe plan risk scores in the insurer rate filing data.  However, we observe all other variables in Equation~\eqref{eqn:ra_formula}, including each plan's risk adjustment transfer, and can therefore solve for the plan risk scores. The plan's utilization share $us_{jmt}(\cdot)$ equals

\vspace{-0.1in}
\begin{equation*}
	us_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) =  \frac{ h_j q_{jmt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\sum_{m \in M, l \in J_{mt}} h_l q_{lmt}(\textit{p},\textit{n};\boldsymbol{\beta})}.
\end{equation*}
A positive difference between the risk share and the utilization share indicates that a plan has high risk relative to its expected utilization and results in the plan receiving a risk adjustment transfer; a negative difference indicates that a plan has low risk relative to its expected utilization and results in the plan paying a risk adjustment transfer.

\paragraph{Calculating Variable Profit:} 
Firms choose premiums and commissions to maximize expected variable profit

\vspace{-0.4in}		
\begin{footnotesize}
\begin{eqnarray}
\label{eqn:profit}
	\pi_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) &=&  R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) -  C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) - \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) + RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) +  RI_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})  - V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})  \nonumber \\
	&=& R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) - (1 - \iota_{ft}) C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) - \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}) + RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})   - V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta}),
\end{eqnarray}
\end{footnotesize}

\noindent where $R_{ft}(\cdot)$ is total premium revenue, $C_{ft}(\cdot)$ is total claims, $RI_{ft}(\cdot)$ is reinsurance received, $V_{ft}(\cdot)$  is variable administrative cost (excluding commissions), and $\iota_{ft}$ is the actuarial value of the reinsurance contract (i.e., the expected percentage of claims paid by the reinsurer).  Total commissions paid $\mathcal{C}_{ft}(\cdot)$ and risk adjustment received  $RA_{ft}(\cdot)$ are as above.

\subsubsection{Stage 2: Household Plan Choice}
Households choose the plan that maximizes their (indirect) utility function

\vspace{-0.4in}
\begin{eqnarray}
\label{eqn:demand_model}
	U_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta}) &\equiv& \beta_{i}^p p_{ijt}(\textit{p}) + \beta_j^{\eta} \eta_{gijt}(\textit{p},\textit{n})   + \beta_i^y y_{ij(t-1)} + x'_{ij}\beta^x + \xi_{j} + \epsilon_{ijt}^d, 
\end{eqnarray}
where $p_{ijt}(\textit{p})$ is household $i$'s (subsidized) premium for plan $j$ in year $t$, $\eta_{gijt}(\textit{p},\textit{n})$ is the commission paid to household $i$'s insurance agent, $y_{ij(t-1)}$ indicates whether household $i$ chose plan $j$ in the previous year, $x_{ij}$ is a vector of observed product characteristics including the plan AV,  $\xi_{j}$ is a vector of unobserved product characteristics, and $\epsilon_{ijt}^d$ is an error term.  CSRs enter utility through the plan AV and premium subsidies reduce the household's premium $p_{ijt}(\textit{p})$.   The utility of the outside option $U_{i0t} = \beta_i^p \rho_{it} + \epsilon_{i0t}$, where $\rho_{it}$ is the household's penalty for not having insurance in year $t$.

The household's premium is calculated as

\vspace{-0.4in}
\begin{eqnarray}
\label{eqn:cons_premium}
	p_{ijt}(\textit{p}) &=& \max \left \lbrace \underbrace{\sigma_{it} p_{jmt}}_{\text{\scriptsize{full premium}}}  - \underbrace{\max\{\sigma_{it} p_{bmt} - \zeta_{it},0\}}_{\text{\scriptsize{premium subsidy}}},0 \right \rbrace, 
\end{eqnarray}
where $\sigma_{it}$ is the household's rating factor as before, $p_{bmt}$ is the base premium of the benchmark plan in market $m$ and year $t$, and $\zeta_{it}$ is the household's income contribution cap.  The household's premium subsidy equals the difference between what the household would pay for the benchmark plan (i.e., $\sigma_{it} p_{bmt}$) and the household's income contribution cap $\zeta_{it}$. The benchmark plan is the second-cheapest silver plan available to the household and varies between consumers because of heterogeneous firm entry. The income contribution cap ranged from $2\%$ of annual income for consumers earning $100\%$ of the federal poverty level (FPL) and $9.5\%$ of annual income for consumers earning $400\%$ of FPL in 2014. The income contribution caps increase slightly each year. Subsidies can be applied to any plan except a catastrophic plan. For some low-income consumers, the premium subsidy may exceed the full premium of certain bronze plans. The subsidy is reduced in these cases to ensure nonnegativity of the premium.

We define the vector of demand parameters $\boldsymbol{\beta} \equiv (\beta_i^p,\beta_j^{\eta},\beta_i^y,\beta^x)$ to accommodate heterogeneous consumer preferences.  The household's premium parameter $\beta_{i}^p \equiv \beta^p + w'_{it} \phi$ varies with household characteristics $w_{it}$.  The household's inertia parameter $\beta_i^y = \beta^y +  x'_{ij} \kappa + w'_{it} \nu$ varies with both household and product characteristics. The steering parameter $\beta_j^{\eta}$ captures the degree to which agent compensation steers  an agent's clients to particular plans.  A high value of $\beta_j^{\eta}$ would indicate that agent compensation is effective in steering consumers.   Because smaller firms offer agents considerably higher commissions, we allow the steering parameter to vary with firm size.

\paragraph{Calculating Demand}
We assume Equation~\eqref{eqn:demand_model}) is a nested logit model such that the vector of error terms $\epsilon_i$ has the generalized extreme value distribution.  The first nest contains all exchange plans and the second nest contains only the outside option.  This two-nest structure captures the key observed substitution pattern between the silver tier and the outside option resulting from the ACA's linkage of CSRs to the purchase of silver plans. Under the assumption that $\epsilon_i$ has the generalized extreme value distribution, the household choice probabilities are

\vspace{-0.4in}		
\begin{eqnarray}
\label{eqn:choice_probs}
	q_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta}) &=&  \frac{e^{V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})/\lambda}\left(\sum_{j } e^{V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})/\lambda}\right)^{\lambda-1}}{1 + \left(\sum_{j} e^{V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})/\lambda}\right)^{\lambda}},
\end{eqnarray}
\noindent where $V_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta}) \equiv \beta_{i}^p p_{ijt}(\textit{p}) + \beta_j^{\eta} \eta_{gijt}(\textit{p},\textit{n})   + \beta_i^y y_{ij(t-1)} + x'_{ij}\beta^x + \xi_{j}$ and $\lambda$ is the nesting parameter. The household choice probabilities in Equation~\eqref{eqn:choice_probs} converge to the standard logit choice probabilities when $\lambda \rightarrow 1$.

The sensitivity of a subsidized consumer's demand $q_{ijt}(\textit{p})$ to a premium change is given by

\vspace{-0.4in}
\begin{eqnarray*}
	\frac{\partial q_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\partial p_{jmt}} &=& \sum_{l \in J_{mt}} \frac{ \partial q_{ijt}(\textit{p},\textit{n};\boldsymbol{\beta})}{\partial p_{ilt}(\textit{p})}\frac{ \partial p_{ilt}(\textit{p})}{\partial p_{jmt}}
\end{eqnarray*}
\noindent for all plans $j,k$ in the set of available plans $J_{mt}$.  Assuming a strictly positive subsidy that does not exceed the full, unsubsidized premium, it follows from Equation~\eqref{eqn:cons_premium} that

\vspace{-0.4in}
\singlespacing
\begin{eqnarray}
\label{eqn:price_partial_formula}
    \frac{\partial p_{ilt}(\textit{p})}{\partial p_{jmt}} = 
    \begin{cases} 
	    0 & l=j, j = b \\		
		\sigma_{it} & l = j, j \neq b \\     
		-\sigma_{it} & l \neq j, j = b \\
		0 & l \neq j, j \neq b  
    \end{cases}.
\end{eqnarray}
\doublespacing
For a non-benchmark plan, an infinitesimal premium increase results in consumers paying more for that plan only.  An infinitesimal increase in the benchmark premium does not affect what subsidized consumers pay for the benchmark plan, but reduces what consumers pay for all other plans because of the larger subsidy. The complex relationship between insurer and consumer premiums, endogenous determination of the benchmark premium, and variation in the benchmark plan across consumers due to heterogeneous entry create significant computational challenges for estimation.  We carefully model the endogenous subsidy design despite the high computational cost because of the critical role premium subsidies play in addressing adverse selection.

\subsubsection*{Equilibrium}

Now we find the first-order necessary conditions for a Nash equilibrium.  Differentiating Equation~\eqref{eqn:profit} with respect to the premium and the commission yields the  first-order conditions

\vspace{-0.4in}	
\begin{footnotesize}
\begin{eqnarray}
\label{eqn:acara_foc_short}
	\frac{\partial R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} + \frac{\partial RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} &=& (1-\iota_{ft}) \frac{\partial C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} + \frac{\partial \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} + \frac{\partial V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial p_{jmt}} \nonumber \\
	\frac{\partial R_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} + \frac{\partial RA_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} &=& (1-\iota_{ft}) \frac{\partial C_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} + \frac{\partial \mathcal{C}_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}} + \frac{\partial V_{ft}(\textit{p},\textit{n};\boldsymbol{\theta})}{\partial n_{gjt}}
\end{eqnarray}
\end{footnotesize}

\noindent for all renewal groups $g$ and markets $m$ in which plan $j$ is offered by the firm in year $t$. In the supplemental appendix, we write every variable in Equations~\eqref{eqn:profit} and \eqref{eqn:acara_foc_short} in terms of three estimable variables: (1) household choice probabilities $q_{ijt}(\textit{p},\textit{n};\boldsymbol{\theta})$; (2) plan risk scores $r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$; and (3) average claims $c_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta})$.  Household choice probabilities are computed using Equation~\eqref{eqn:choice_probs}.  We calculate plan risk scores as a function of observable enrollee characteristics and the plan generosity using the estimating equation,

\vspace{-0.4in}
\begin{eqnarray}
\label{eqn:risk_score_estimation}
	\ln r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) &=& \sum_{d \in D} \gamma^d s_{djmt}(\textit{p},\textit{n};\boldsymbol{\theta}) + MT'_j \gamma^{MT} + \epsilon_{jmt}^r.
\end{eqnarray}
The predicted demographic share $s_{djmt}(\cdot)$ is the share of plan $j$'s enrollment in market $m$ and year $t$ with demographic characteristic $d$, $MT_j$ is a vector metal tier fixed effects,  $\epsilon_{jmt}^r$ is an error term, and the vector of risk score parameters $\boldsymbol{\gamma} = (\gamma^d,\gamma^{MT},\gamma^{n})$. The demographic shares are computed by aggregating the household choice probabilities.   We calculate plan average claims as a function of the plan risk score using the estimating equation

\vspace{-0.4in}	
\begin{eqnarray}
\label{eqn:avg_claims_estimation}
	\ln c_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) &=& \mu^r \ln r_{jmt}(\textit{p},\textit{n};\boldsymbol{\theta}) + x'_j  \mu^{x}  + \mu^l l_t +   n'_m \mu^{n} + \epsilon_{jmt}^c,
\end{eqnarray}
where $r_{jmt}(\cdot)$ is the predicted risk score computed using Equation~\eqref{eqn:risk_score_estimation}, $x_j$ are product characteristics (not including plan AV), $l_t$ is a linear trend, $n'_m$ are market fixed effects, $\epsilon_{jmt}^c$ is an error term, and  $\boldsymbol{\mu} = (\mu^r,\mu^x,\mu^l,\mu^n)$ are the claims parameters.

\subsection{Estimation}
\label{subsec:steering-methods}

In this section, we explain how we use the generalized method of moments (GMM) to estimate the parameter vector $\boldsymbol{\theta}$.  We construct four sets of moment conditions: (1) demand moments that match observed choices and predicted household choice probabilities; (2) risk score moments that match observed and predicted plan risk scores; and (3) average claims moments that match observed and predicted plan average claims; and (4) the first-order conditions for profit maximization in Equation~\eqref{eqn:acara_foc_short}. These moment conditions are summarized below:
